\bf Supervised Tensor Decomposition with Interactive Side Information

### 2020-09-22 tensor_regress/JCGS_submission/jcgs-tensor/supp.tex
jcgs
update

\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{-.3in}%
\addtolength{\topmargin}{-.8in}%

@@

\usepackage{xr}
\externaldocument{regress}

@@

\section{Proofs}\label{sec:appedix}
We restate the Theorem 4.1 from the main text. 
 \begin{thm}[Statistical convergence]\label{thm:main}
Consider a data tensor generated from model~\eqref{eq:decomp}, where the entries are conditionally independent realizations from an exponential family. Let $(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)$ be the M-estimator in~\eqref{eq:MLE} and $\hat \tB=\hat \tC\times \hat \mM_1\times\cdots \times \hat \mM_K$. Define $r_\textup{total}=\prod_k r_k$ and $r_{\max}=\max_k r_k$. Under Assumptions A1 and A2 with scaled feature matrices $\check \mX_k= \sqrt{d_k}\mX_k$, or under Assumptions A1' and A2 with original feature matrices, there exist two positive constants $C_1=C_1(\alpha,K), C_2=C_2(\alpha, K)>0$ independent of dimensions $\{d_k\}$ and $\{p_k\}$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$, 
\begin{equation}\label{eq:bound}
    \FnormSize{}{\trueB- \hat \tB}^2\leq \frac{ C_{2} r_{\textup{total}}}{r_{\max}}\frac{\sum_k p_k}{\prod_k d_k}.
\end{equation}
Furthermore, if the unfolded core tensor has non-degenerate singular values at mode $k\in[K]$, i.e., $\sigma_{\min}(\textup{Unfold}_k(\tC_\textup{true})) \geq c>0$ for some constant $c$, then
\begin{equation}\label{eq:sinebound}
\textup{sin}^2 \Theta(\trueM,\ \hat \mM_k) \leq  \frac{ C_{2}r_{\textup{total}}}{ r_{\max}\sigma^2_{\min}(\textup{Unfold}_k(\tC_\textup{true}))}\frac{\sum_k p_k}{\prod_k d_k}.
\end{equation}
\end{thm}

@@

\begin{proof}[Proof of Theorem~\ref{thm:KL}]
The proof is similar to~\citet{pmlr-v108-berthet20a}. We sketch the main steps here for completeness. Recall that $\ell(\tB)=\mathbb{E}(\tL_{\tY}(\tB))$. By the definition of KL divergence, we have that,
\begin{align}
\ell(\hat \tB)&=\ell(\trueB)-\sum_{(i_1,\ldots,i_K)} KL(\theta_{\text{true}, i_1,\ldots, i_K}, \hat \theta_{i_1,\ldots,i_K})\\
&=\ell(\trueB)-\text{KL}(\mathbb{P}_{\tY_{\text{true}}},\ \mathbb{P}_{\hat \tY}),
\end{align}
where $\mathbb{P}_{\tY_{\text{true}}}$ denotes the distribution of $\tY|\tX$ with true parameter $\trueB$, and $\mathbb{P}_{\hat \tY}$ denotes the distribution with estimated parameter $\hat \tB$. Therefore
\begin{align}
\text{KL}(\mathbb{P}_{\tY_{\text{true}}},\ \mathbb{P}_{\hat \tY}) &= \ell(\trueB)-\ell(\hat \tB)\\
&={1\over 2}\sum_{i_1\ldots,i_K}b''(\check \theta_{i_1,\ldots,i_K})(\theta_{i_1,\ldots,i_K}-\theta_{\text{true},i_1,\ldots,i_K})^2\\
&\leq {U\over 2} \FnormSize{}{\Theta-\trueT}^2\\
&\leq C_4{\prod_k r_k \over \max r_k}\sum_k p_k,
\end{align}
where the second line comes from~\eqref{eq:log}, and the last line is derived from~\eqref{eq:theta}. Notice that $C_4 = C(\alpha,K,U,c_1,c_2)>0$ in Assumption 1 and $C_4(\alpha,K,U,\gamma)>0$ in Assumption 1' are constants that do not depend on the dimension $\{d_k\}$ and $\{p_k\}$. 
\end{proof}

\section{Algorithm properties}\label{sec:SAlgorithm}
In this section, we provide the convergence properties of Algorithm~\ref{alg:B}.
For notational convenience, we drop the subscript $\tY$ from the objective $\tL_\tY(\cdot)$ and simply write as $\tL(\cdot)$. Let $\tA = (\tC, \mM_1,\ldots,\mM_K) \in \mathbb{R}^{d_{\text{total}}}$ denote the collection of decision variables in the alternating optimization, where $d_{\text{total}} = \prod_k r_k + \sum_k r_kd_k$. We introduce the equivalent relationship induced by orthogonal transformation. Let $\mathbb{O}_{d,r}$ be the collection of all $d$-by-$r$ matrices with orthogonal columns, $\mathbb{O}_{d,r}:=\{\mP\in\mathbb{R}^{d\times r}\colon \mP^T\mP = \mathbf{1}_r\}$, where $\mathbf{1}_r$ is the $r$-by-$r$ identity matrix.

\begin{defn}[Equivalance relation]\label{eq:equivalance}
Two parameters $\tA'=(\tC', \mM'_1,\ldots,\mM'_k)$ and $\tA=(\tC, \mM_1,\ldots,\mM_k)$ are called equivalent, denoted $\tA\sim \tA'$, if and only if there exist a set of orthogonal matrices $\mP_k \in \mathbb{O}_{d_k,r_k}$ such that
\begin{align}
	\mM'_k \mP^T_k = \mM_k ,\ \forall k \in [K], \quad \text{and}\quad \tC' \times_1 \mP_1 \times_2  \dots \times_K \mP_K = \tC.
\end{align}
\end{defn}
Equivalently, two parameters $\tA,\ \tA'$ are equivalent if the corresponding Tucker tensors are the same, $\tB(\tA)=\tB'(\tA')$.

\section{Computational complexity}
The computational complexity of our Algorithm~\eqref{alg:B} is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the data tensor. More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor. The update of each factor matrix $\mM_k$ involves a GLM with a $d$-length response, and $d$-by-$(r_kp_k)$ feature matrix. Solving such a GLM requires $O(dr^3_kp^3_k)$, and therefore the cost for updating $K$ factors in total is $O( d\sum_k r^3_k p_k^3)$. This complexity in tensor dimension matches with the classical tensor decomposition~\citep{kolda2009tensor}.

>>>>

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

@@

\usepackage{xr}
\externaldocument{jcgs-tensor}

@@

\section{Proofs}\label{sec:appedix}

\subsection{Proofs of Proposition~\ref{prop:alg}}\label{sec:SAlgorithm}
For notational convenience, we drop the subscript $\tY$ from the objective $\tL_\tY(\cdot)$ and simply write as $\tL(\cdot)$. Let $\tA = (\tC, \mM_1,\ldots,\mM_K) \in \mathbb{R}^{d_{\text{total}}}$ denote the collection of decision variables in the alternating optimization, where $d_{\text{total}} = \prod_k r_k + \sum_k r_kd_k$. We recall the equivalent relationship induced by orthogonal transformation.

\begin{defn}[Equivalence class]
Two parameters $\tA'=(\tC', \mM'_1,\ldots,\mM'_k)$ and $\tA=(\tC, \mM_1,\ldots,\mM_k)$ are called equivalent, denoted $\tA\sim \tA'$, if and only if there exist a set of orthogonal matrices $\mP_k \in \mathbb{O}_{d_k,r_k}$ such that
\begin{align}\label{eq:equivalance}
	\mM'_k \mP^T_k = \mM_k ,\ \forall k \in [K], \quad \text{and}\quad \tC' \times_1 \mP_1 \times_2  \dots \times_K \mP_K = \tC.
\end{align}
\end{defn}

\subsection{Proof of Theorem 4.1}
We restate the Theorem 4.1 from the main text. 
 \begin{thm}[Statistical convergence]\label{thm:main}
Consider a data tensor generated from supervised tensor model, where the entries are conditionally independent realizations from an exponential family. Let $(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)$ be the M-estimator and $\hat \tB=\hat \tC\times \hat \mM_1\times\cdots \times \hat \mM_K$. Define $r_\textup{total}=\prod_k r_k$ and $r_{\max}=\max_k r_k$. Under Assumptions A1 and A2 with scaled feature matrices $\check \mX_k= \sqrt{d_k}\mX_k$, or under Assumptions A1' and A2 with original feature matrices, there exist two positive constants $C_1=C_1(\alpha,K), C_2=C_2(\alpha, K)>0$ independent of dimensions $\{d_k\}$ and $\{p_k\}$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$, 
\begin{equation}\label{eq:bound}
    \FnormSize{}{\trueB- \hat \tB}^2\leq \frac{ C_{2} r_{\textup{total}}}{r_{\max}}\frac{\sum_k p_k}{\prod_k d_k}.
\end{equation}
Furthermore, if the unfolded core tensor has non-degenerate singular values at mode $k\in[K]$, i.e., $\sigma_{\min}(\textup{Unfold}_k(\tC_\textup{true})) \geq c>0$ for some constant $c$, then
\begin{equation}\label{eq:sinebound}
\textup{sin}^2 \Theta(\trueM,\ \hat \mM_k) \leq  \frac{ C_{2}r_{\textup{total}}}{ r_{\max}\sigma^2_{\min}(\textup{Unfold}_k(\tC_\textup{true}))}\frac{\sum_k p_k}{\prod_k d_k}.
\end{equation}
\end{thm}

@@

\subsection{Proof of Corollary~\ref{thm:KL}}
\begin{proof}[Proof of Corollary~\ref{thm:KL}]
The proof is similar to~\citet{pmlr-v108-berthet20a}. We sketch the main steps here for completeness. Recall that $\ell(\tB)=\mathbb{E}(\tL_{\tY}(\tB))$. By the definition of KL divergence, we have that,
\begin{align}
\ell(\hat \tB)&=\ell(\trueB)-\sum_{(i_1,\ldots,i_K)} KL(\theta_{\text{true}, i_1,\ldots, i_K}, \hat \theta_{i_1,\ldots,i_K})\\
&=\ell(\trueB)-\text{KL}(\mathbb{P}_{\tY_{\text{true}}},\ \mathbb{P}_{\hat \tY}),
\end{align}
where $\mathbb{P}_{\tY_{\text{true}}}$ denotes the distribution of $\tY|\tX$ with true parameter $\trueB$, and $\mathbb{P}_{\hat \tY}$ denotes the distribution with estimated parameter $\hat \tB$. Therefore
\begin{align}
\text{KL}(\mathbb{P}_{\tY_{\text{true}}},\ \mathbb{P}_{\hat \tY}) &= \ell(\trueB)-\ell(\hat \tB)\\
&={1\over 2}\sum_{i_1\ldots,i_K}b''(\check \theta_{i_1,\ldots,i_K})(\theta_{i_1,\ldots,i_K}-\theta_{\text{true},i_1,\ldots,i_K})^2\\
&\leq {U\over 2} \FnormSize{}{\Theta-\trueT}^2\\
&\leq C_4{\prod_k r_k \over \max r_k}\sum_k p_k,
\end{align}
where the second line comes from~\eqref{eq:log}, and the last line is derived from~\eqref{eq:theta}. Notice that $C_4 = C(\alpha,K,U,c_1,c_2)>0$ in Assumption 1 and $C_4(\alpha,K,U,\gamma)>0$ in Assumption 1' are constants that do not depend on the dimension $\{d_k\}$ and $\{p_k\}$. 
\end{proof}
